---
title: "feature generation and model"
author: "XU ZHEN"
date: "2022-11-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr) 
library(readxl)
library(ggplot2)
#install.packages("ggpubr")
library(ggpubr)
```

## something we should notice
1. level of measure: nominal level, ordinal level, interval level, ratio level
2. standardized


## Reading data

```{r pressure}
df_test_clean<-read.csv("output data/df_test_clean.csv")
df_train_clean<-read.csv("output data/df_train_clean.csv")
```


#### level of measure: nominal level, ordinal level, interval level, ratio level, scholarship

nominal level: cohort.term,Marital.Status

ordinal level: Father.s.Highest.Grade.Level, Mother.s.Highest.Grade.Level,

interval level:

ratio level: Adjusted.Gross.Income, Parent.Adjusted.Gross.Income


## finance
#### cohort.term, Marital.Status, Father.s.Highest.Grade.Level, Mother.s.Highest.Grade.Level
```{r}
df_t<-df_train_clean

df_t$Dropout = factor(df_t$Dropout,levels = c(0,1),labels = c("Grad", "dropout"))

################################   cohort.term  ################################ 
perc<-ggplot(data = df_t, mapping = aes(x = cohort.term, fill =Dropout)) + 
ylab("perc") + geom_bar(position="fill",alpha = 0.3)
count<-ggplot(df_t, mapping=aes(x = cohort.term, fill = Dropout)) +geom_bar(alpha = 0.3)
ggarrange(perc, count, labels = c("perc", "count"),ncol = 1, nrow = 2)

################################   Marital.Status  ################################ 
perc<-ggplot(data = df_t, mapping = aes(x = Marital.Status, fill =Dropout)) + 
ylab("perc") + geom_bar(position="fill",alpha = 0.3)
count<-ggplot(df_t, mapping=aes(x = Marital.Status, fill = Dropout)) +geom_bar(alpha = 0.3)
ggarrange(perc, count, labels = c("perc", "count"),ncol = 1, nrow = 2)

################################   Father.s.Highest.Grade.Leve  ################################ 
perc<-ggplot(data = df_t, mapping = aes(x = Father.s.Highest.Grade.Level, fill =Dropout)) + 
ylab("perc") + geom_bar(position="fill",alpha = 0.3)
count<-ggplot(df_t, mapping=aes(x = Father.s.Highest.Grade.Level, fill = Dropout)) +geom_bar(alpha = 0.3)
ggarrange(perc, count, labels = c("perc", "count"),ncol = 1, nrow = 2)

################################   Mother.s.Highest.Grade.Level  ################################ 
perc<-ggplot(data = df_t, mapping = aes(x = Mother.s.Highest.Grade.Level, fill =Dropout)) + 
ylab("perc") + geom_bar(position="fill",alpha = 0.3)
count<-ggplot(df_t, mapping=aes(x = Mother.s.Highest.Grade.Level, fill = Dropout)) +geom_bar(alpha = 0.3)
ggarrange(perc, count, labels = c("perc", "count"),ncol = 1, nrow = 2)

```

cohort term 可以放进去看一下！

```{r}
###############  Adjusted.Gross.Income, Parent.Adjusted.Gross.Income  ################################ 
#######需要去掉离群点
ggplot(df_t, aes(x = Adjusted.Gross.Income)) +geom_boxplot()
ggplot(df_t, aes(x = log10(Adjusted.Gross.Income), fill = Dropout)) +geom_density(alpha = 0.3)
#####标准化一下做相关性分析，检验正态分布，

###normal distribution test

###### add column
df_t$Adjusted.Gross.Income<-abs(df_t$Adjusted.Gross.Income)
df_t$Parent.Adjusted.Gross.Income<-abs(df_t$Parent.Adjusted.Gross.Income)

ks.test(scale(df_t$Adjusted.Gross.Income),"pnorm")
qqnorm(df_t$Adjusted.Gross.Income)

ks.test(scale(df_t$Parent.Adjusted.Gross.Income),"pnorm")
qqnorm(df_t$Parent.Adjusted.Gross.Income)

###### add column
df_t$sum_income<-df_t$Adjusted.Gross.Income+df_t$Parent.Adjusted.Gross.Income

ks.test(scale(df_t$sum_income),"pnorm")
qqnorm(df_t$sum_income)

####  Mann-Whitney U test
wilcox.test(df_t$Parent.Adjusted.Gross.Income~df_t$Dropout)
wilcox.test(df_t$Adjusted.Gross.Income~df_t$Dropout)
wilcox.test(df_t$sum_income~df_t$Dropout)


```
三个差异均显著，保留一个


```{r}

################################   scholarship     ################################ 
##### add column

for (i in 1:nrow(df_t)){
  df_t$total_Loan[i]<-sum(df_t[i,grep("\\.Loan", colnames(df_t))])
}

for (i in 1:nrow(df_t)){
  df_t$total_Scholarship[i]<-sum(df_t[i,grep("\\.Scholarship", colnames(df_t))])
}

for (i in 1:nrow(df_t)){
  df_t$total_Work_Study[i]<-sum(df_t[i,grep("\\.Work\\.Study", colnames(df_t))])
}

for (i in 1:nrow(df_t)){
  df_t$total_Grant[i]<-sum(df_t[i,grep("\\.Grant", colnames(df_t))])
}



##### plot
ggplot(df_t, aes(x = log10(total_Loan), fill = Dropout)) +geom_density(alpha = 0.3)
ggplot(df_t, aes(x = log10(total_Scholarship), fill = Dropout)) +geom_density(alpha = 0.3)
ggplot(df_t, aes(x = log10(total_Work_Study), fill = Dropout)) +geom_density(alpha = 0.3)
ggplot(df_t, aes(x = log10(total_Grant), fill = Dropout)) +geom_density(alpha = 0.3)


ggplot(data = df_t, mapping = aes(x = log10(total_Loan), fill =Dropout)) + ylab("perc") + geom_histogram(position="fill",alpha = 0.3)

ggplot(data = df_t, mapping = aes(x = total_Scholarship, fill =Dropout)) + ylab("perc") + geom_histogram(position="fill",alpha = 0.3)

ggplot(data = df_t, mapping = aes(x = total_Work_Study, fill =Dropout)) + ylab("perc") + geom_histogram(position="fill",alpha = 0.3)

ggplot(data = df_t, mapping = aes(x = total_Grant, fill =Dropout)) + ylab("perc") + geom_histogram(position="fill",alpha = 0.3)

############  取一下log !!!!!

ks.test(scale(df_t$total_Loan),"pnorm")
qqnorm(df_t$total_Loan)

####  Mann-Whitney U test
wilcox.test(df_t$total_Loan~df_t$Dropout)
wilcox.test(df_t$total_Scholarship~df_t$Dropout)
wilcox.test(df_t$total_Work_Study~df_t$Dropout)
wilcox.test(df_t$total_Grant~df_t$Dropout)

```

## static


State"                          "RegistrationDate"              
 [37] "Gender"                         "BirthYear"                     
 [39] "BirthMonth"                     "Hispanic"                      
 [41] "AmericanIndian"                 "Asian"                         
 [43] "Black"                          "NativeHawaiian"                
 [45] "White"                          "TwoOrMoreRace"                 
 [47] "HSDip"                          "HSDipYr"                       
 [49] "HSGPAUnwtd"                     "EnrollmentStatus"              
 [51] "NumColCredAttemptTransfer"      "NumColCredAcceptTransfer"      
 [53] "CumLoanAtEntry"                 "HighDeg"                       
 [55] "MathPlacement"                  "EngPlacement"                  
 [57] "GatewayMathStatus"              "GatewayEnglishStatus"     
```{r}

################################   gender  ################################ 
perc<-ggplot(data = df_t, mapping = aes(x = Gender, fill =Dropout)) + 
ylab("perc") + geom_bar(position="fill",alpha = 0.3)
count<-ggplot(df_t, mapping=aes(x = Gender, fill = Dropout)) +geom_bar(alpha = 0.3)
ggarrange(perc, count, labels = c("perc", "count"),ncol = 1, nrow = 2)

################################   race  ################################ #
for (i in 1:nrow(df_t)){
  if (df_t$Asian[i]==1 |df_t$White[i]==1){
    df_t$race[i]<-1
  }
  else 
    df_t$race[i]<-0
}
by(df_t$Dropout,df_t$race,summary)

```
## progress

```{r}
for (i in 1:nrow(df_t)){
  df_t[i,"cohort_year"]<-substring((df_t[i,"cohort"]),1,4)
}

for (i in 1:nrow(df_t)){
  if (df_t$Asian[i]==1 |df_t$White[i]==1){
    df_t$race[i]<-1
  }
  else 
    df_t$race[i]<-0
}

# ###### first term Major
col_Major1<-grep("Major1", colnames(df_t))
col_Major1
colnames(df_t)
df_t[,col_Major1][is.na(df_t[,col_Major1])] <- "0"

for (i in 1:nrow(df_t)){
  if (df_t[i,"cohort.term"]==1){
    df_t[i,"first_term_Major"]<-df_t[i,paste("Major1","Fall",df_t[i,"cohort_year"],sep="_")]
  }
  else if(df_t[i,"cohort.term"]==3){
    if(df_t[i,paste("Major1","Fall",df_t[i,"cohort_year"],sep="_")]>0){
      df_t[i,"first_term_Major"]<-df_t[i,paste("Major1","Fall",df_t[i,"cohort_year"],sep="_")]
    }
    else
      df_t[i,"first_term_Major"]<-df_t[i,paste("Major1","Spring",as.numeric(df_t[i,"cohort_year"])+1,sep="_")]
  }
}

```

